---
title: "Recitation 11: Weighting"
author: "Seamus Wagner"
date: "November 9, 2021"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = "center", echo=T, warning=FALSE, message=FALSE)
```

```{r, echo = F}
library(dplyr)
library(data.table)
library(ggplot2)
```

# Lab preview

Since we had to cruise through the code demo, I will offer a bit of review focused on the problem set. My goal here is to do a bit of a guided walkthrough where we fill out code and I will update the .rmd after recitation with what we come up with. That way we can have a bit of a more engaged session and you still get the benefit of having code to refer to along with comments for the problem set. I will be focusing on the questions that are more code-intensive and will leave the others to clarifying questions after you have spent some time with the recitation. 

Problem 1
```{r}
rm(list=ls())
set.seed(123321)
library(data.table)
n_covariates <- 8
n_obs <- 1000
# How to approach question 1?

# Why does this only give 1 column, how to make it 8?
x <- matrix(rnorm(n_obs * n_covariates))

# Once you have it for 8, how to make a data.table with the other three variables?
#dt[, ':=' (
#id = ?,
#Y1 = ?,
#Y0 = ?
#)]
#ps. use ctrl+shift+a to reorganize chunks of code you highlight, try it on above once you uncomment it and fill it out.
#pps. if you hold alt+up/down arrow, you can shift a line of code up or down and swap it with the existing line, try switching Y1 and Y0 so that Y0 appears first.
```

Problem 2

```{r}
# Once we have some data, how do we set up the logit model? 
# logitmodel <- lm()

# After, how do we calculate e? If lost, check the weighting demo propensity score sections. 
# ??
```

Problem 4
```{r}
# We will go over what the function is and leave 5 and 6 to interpretation for you to work on alone or in groups
```

Problem 7

```{r}
# What is the data.table command to filter
```

Problem 8

```{r}
# Outer performs an operation on two arrays with syntax x,y, "function").
```

Problem 9

```{r}
# How would we set this up? What functions give us the min?
```

Problem 10

```{r}
# Break it into two parts, making the data, then graphing. Merging, filtering, iteration, or rbind/cbind are options here.
```

Problem 11

```{r}
# What kind of merge would give us this? What alternatives are there? 
```



